Constraints on quasi--normal-mode frequencies with LIGO-Virgo binary--black-hole observations

### 2021-05-03 papers/intro_paper/intro_paper.tex
small changes

The observation of GWs has had significant astrophysical and cosmological
implications~\cite{TheLIGOScientific:2016htt,GBM:2017lvd,Monitor:2017mdv,Abbott:2017xzu}. It
has also allowed us to probe fundamental
physics and test predictions of Einstein's theory of General Relativity
(GR) in the previously unexplored highly-dynamical, and strong field
regime~\cite{TheLIGOScientific:2016src,Abbott:2018lct,LIGOScientific:2019fpa,Abbott:2020jks}. In GR, a binary black hole (BBH) system is described by
three distinct phases: an early \textit{inspiral}~\cite{Blanchet:2013haa}, where the two
compact objects spiral in loosing energy because of the emission of GWs, a \textit{merger}~\cite{Pretorius:2005gq,Campanelli:2005dd,Baker:2005vv} marked by the
formation of a common apparent horizon, and a \textit{ringdown}~\cite{Vishveshwara:1970zz,Vishveshwara:1970cc,Press:1971wr,Chandrasekhar:1975zza,Detweiler:1980gk}, during which the newly formed remnant object settles down to a Kerr BH emitting quasi-normal-modes (QNMs) (i.e., damped oscillations with specific, discrete frequencies and decay times).

@@

We now demonstrate our method using synthetic-signal injections describing GWs
from BBHs in GR. We employ coloured Gaussian noise with PSDs for LIGO and
Virgo detectors during the fourth observing (O4) run (i.e., at design sensitivity), which 
is expected to start in the second half of 2022~\cite{AdvLIGOPSD,TheVirgo:2014hva}.
For the mock BBH signals, we choose parameters similar to two specific GW events, GW150914~\cite{Abbott:2016blz} and
GW190521~\cite{Abbott:2020tfl}. We list them in Table~\ref{tab:injection_values}.
These two binary systems are representative of the kind of systems for which
the QNM measurement is most suitable, notably high-mass BBH events which are loud enough that the
pre- and post-merger SNRs return reliable parameter-estimation results.

@@

Furthermore, when assuming that the deviations from GR do not vary appreciably over the GW events 
that we have analyzed, and combine the likelihood functions, we have obtained the most stringent bounds  
on the dominant QNM $(\df{220}=0.02^{+0.04}_{-0.04}, \dtau{220}=0.10^{+0.14}_{-0.14})$. Those 
constraints are compatible, but sligthly better (especially for the decay time) than the ones recently reported in Ref.~\cite{Carullo:2021dui} 
(see first row in Table II therein), where QNM frequencies were inferred using only the post-merger part of 
the signal. We note that Ref.~\cite{Carullo:2021dui} used a larger set of GW events from O1, O2 and O3a than we did, although we expect 
that the extra GW events will not contribute significantly to the combined bound since they have lower SNRs.

>>>>

The observation of GWs has had significant astrophysical and cosmological
implications~\cite{TheLIGOScientific:2016htt,GBM:2017lvd,Monitor:2017mdv,Abbott:2017xzu}. It
has also allowed us to probe fundamental
physics and test predictions of Einstein's theory of General Relativity
(GR) in the previously unexplored highly-dynamical, and strong field
regime~\cite{TheLIGOScientific:2016src,Abbott:2018lct,LIGOScientific:2019fpa,Abbott:2020jks}. In GR, a binary black hole (BBH) system is described by
three distinct phases: an early \textit{inspiral}~\cite{Blanchet:2013haa}, where the two
compact objects spiral in losing energy because of the emission of GWs, a \textit{merger}~\cite{Pretorius:2005gq,Campanelli:2005dd,Baker:2005vv} marked by the
formation of a common apparent horizon, and a \textit{ringdown}~\cite{Vishveshwara:1970zz,Vishveshwara:1970cc,Press:1971wr,Chandrasekhar:1975zza,Detweiler:1980gk}, during which the newly formed remnant object settles down to a Kerr BH emitting quasi-normal-modes (QNMs) (i.e., damped oscillations with specific, discrete frequencies and decay times).

@@

We now demonstrate our method using synthetic-signal injections describing GWs
from BBHs in GR. We employ coloured Gaussian noise with PSDs for LIGO and
Virgo detectors during the fourth observing (O4) run~\cite{Abbott:2020qfu}, which 
is expected to start in the second half of 2022~\cite{AdvLIGOPSD,TheVirgo:2014hva}.
For the mock BBH signals, we choose parameters similar to two specific GW events, GW150914~\cite{Abbott:2016blz} and
GW190521~\cite{Abbott:2020tfl}. We list them in Table~\ref{tab:injection_values}.
These two binary systems are representative of the kind of systems for which
the QNM measurement is most suitable, notably high-mass BBH events which are loud enough that the
pre- and post-merger SNRs return reliable parameter-estimation results.

@@

Furthermore, when assuming that the deviations from GR do not vary appreciably over the GW events 
that we have analyzed, and combine the likelihood functions, we have obtained the most stringent bounds  
on the dominant QNM $(\df{220}=0.02^{+0.04}_{-0.04}, \dtau{220}=0.10^{+0.14}_{-0.14})$. Those 
constraints are compatible, but slightly better (especially for the decay time) than the ones recently reported in Ref.~\cite{Carullo:2021dui} 
(see first row in Table II therein), where QNM frequencies were inferred using only the post-merger part of 
the signal. We note that Ref.~\cite{Carullo:2021dui} used a larger set of GW events from O1, O2 and O3a than we did, although we expect 
that the extra GW events will not contribute significantly to the combined bound since they have lower SNRs.
